Species motif participation provides unique information about species risk of extinction

### 2022-06-02 manuscript/reorder_main_text.tex
main text revisions per R1

    %\subsection*{2 Global and local network properties are known to affect stability}
    How disturbances, such as population decline or extinction, lead to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
    Several global (whole-network) as well as local (single-species) properties are known to influence how species respond to disturbances. Two well-known, influential global properties are network size, the number of species present in the network, and network connectance, the number of realized feeding interactions.
    Based on simulations of random networks, large networks were initially thought to be less stable~\citep{May1972}.
    The presence of large, stable, empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
    Highly-connected networks may be more resistant to secondary extinctions \citep{Dunne2002, Eklof2006} because consumers, on average, have access to a larger number of prey species. However, in a well-connected network a disturbance may spread more easily and cause more secondary extinctions due to the presence of many possible pathways \citep{}.
    The net effect of connectance on extinction therefore likely depends on the severity of disturbance and the ability of consumers to subsist on a reduced set of prey.

    Local properties known to affect species' response to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator). 
    Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, including secondary extinction due to reduced plant populations.

    However, in-degree and trophic level are both relatively coarse descriptions of a species' role in its food web~\citep{Cirtwill2018FoodWebs}. 
    As well as the properties of the consumer itself, a consumer's risk of secondary extinction after a disturbance to plants depends on the extinction risk of its prey.
    For example, a generalist consuming many specialist (and therefore high-risk) prey may, for example, be at greater risk of extinction than a specialist which consumes a highly generalist resource that is very unlikely to go extinct.

@@

    While the potential relationships between motif participation and species-level extinction risk have not yet been tested to our knowledge, it has been shown that the frequencies of different motifs within a whole network are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}. 
    Given these relationships between network-level motif frequencies and overall community stability, it is worth investigating whether a species' participation in different motifs also affects its extinction risk.
    We are also interested in whether the relationships between network structure and extinction risk depend on the severity of disturbance experienced by the community.

    Naturally, a larger disturbance tends to lead to more secondary extinctions \citep{}.
    Disturbances to plant species can also affect some consumers more strongly than others  \citep{byrnes2011climate}. 
    In empirical studies of kelp forests, the magnitude of the disturbances to basal resources can alter the competition between species in the benthic community \citep{detmer2021variation} and change the percent coverage of different taxonomic groups \citep{carnell2020more}.

@@

	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and STL), 3) how network mean persistence (the average persistence of consumers) varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties (S, C, level of disturbance, and their interactions).
	See Appendix \emph{3.1 - 3.5} for details. 
	All analyses are briefly described below and were performed using the R~\citep{R} packages \emph{lmerTest}~\citep{lmerTest}, \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan}.

@@

>>>>

    %\subsection*{2 Global and local network properties are known to affect stability}
    How disturbances, such as population decline or extinction, lead to loss of other species via direct and indirect effects (i.e., secondary extinctions) is a vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
    Several global (whole-network) as well as local (single-species) properties are known to influence how species respond to disturbances. Two well-known, influential global properties are network size, the number of species present in the network, and network connectance, the number of realized feeding interactions.
    Based on simulations of random networks, large networks were initially thought to be less stable~\citep{May1972}.
    The presence of large, stable, empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
    Highly-connected networks may be more resistant to secondary extinctions because consumers, on average, have access to a larger number of prey species and can switch from a declining or extinct prey to another resource~\citep{Dunne2002, Eklof2006,Baumgartner2020}. However, a well-connected network also includes more pathways through which a disturbance can spread~\citep{Vieira2015}.
    Depending on the strength and arrangements of interactions, this means that species may be impacted through several paths, amplifying the original disturbance and causing more secondary extinctions.
    The net effect of connectance on extinction therefore likely depends on finer-scale network structure, as well as the type and severity of disturbance.

    Local properties known to affect species' response to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator). 
    Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, including secondary extinction due to reduced plant populations.
    However, in-degree and trophic level are both relatively coarse descriptions of a species' role in its food web~\citep{Cirtwill2018FoodWebs}. 
    As well as the properties of the consumer itself, a consumer's risk of secondary extinction after a disturbance to plants depends on the extinction risk of its prey.
    For example, a generalist consuming many specialist (and therefore high-risk) prey may, for example, be at greater risk of extinction than a specialist which consumes a highly generalist resource that is very unlikely to go extinct.

@@

    While the potential relationships between motif participation and species-level extinction risk have not yet been tested to our knowledge, it has been shown that the frequencies of different motifs within a whole network are associated with community stability \citep{prill2005dynamic, bascompte2005simple} and with changes to the composition of plant communities \citep{giling2019plant}. 
    These relationships between network-level motif frequencies (`motif profiles') and overall community stability mean that a network's motif profile could affect the average extinction risk for consumers within the network.
    Moreover, if network motif profiles are related to average extinction risk, it is worth investigating whether a species' participation in different motifs also affects its extinction risk.

    As mentioned above, the severity of a disturbance may interact with network structure to affect secondary extinctions.
    A mild disturbance (e.g., a small increase in the probability of basal resources going extinct) is intuitively likely to cause fewer secondary extinctions than a severe disturbance, and the difference in effects can depend on network structure~\citep{Baumgartner2015}.
    Empirical studies have shown that disturbances to plant species can also affect some consumers more strongly than others  \citep{byrnes2011climate}, and that the strength of disturbance also affects which consumer species are most strongly affected~\citep{detmer2021variation,carnell2020more}. 
    By taking advantage of recent advances in rapid simulation of ecological networks, we are able to simulate a range of disturbance intensities and test whether meso-scale network structure also interacts with disturbance.

@@

    	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and STL), 3) how network mean persistence (the average persistence of consumers) varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties (S, C, in-degree, and trophic level).
        Where applicable, we also consider how interactions between network structure and disturbance level might affect extinction risk.
    	See Appendix \emph{3.1 - 3.5} for details. 
    	All analyses are briefly described below and were performed using the R~\citep{R} packages \emph{lmerTest}~\citep{lmerTest}, \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan}.

@@

    \begin{figure}[hb!]
        \centering
        \includegraphics[width=0.85\textwidth]{figures/persistence_vs_motifs.eps}
        \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis). The effect of participating in each motif is based on the fixed effects in Equation~\ref{propreq}. The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance). 95\% confidence intervals for each line are shown in grey. Note that omnivory made up a smaller proportion of species' roles than other motifs; lines are plotted over the observed ranges of motif participation.}
    \label{fig:prop_lmer_all}
    \end{figure}
